Supporting Information

### 2017-04-24 code/rmarkdown/article.Rmd
assorted fixes

Abstract word count: 347 / 350
Total word count: XXXX / 7000
Number of references: 42
Number of figures: 8
Number of tables: 0
Number of textboxes: 0
Number of equations: 4
Short running title:
Short abstract: Prioritize biodiversity processes and patterns using raptr https://github.com/jeffreyhanson/raptr
Keywords: conservation, biodiversity, mixed integer linear programming, optimization, protected areas

@@

1. An underlying aim in conservation planning is to maximize the long-term persistence of biodiversity. To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved. One way to conserve such processes at the feature level (eg. species, ecosystem) is to preserve a sample of the feature (eg. individuals, areas) that is representative in terms of the intrinsic or extrinsic physical attributes that underpin the process of interest. For example, by conserving a sample of populations with local adaptations that is representative in terms of the full variation found in the species--physical attributes associated with adaptation--protected areas can foster adaptive processes by ensuring these adaptations are not lost. Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of area or habitat for each feature. Little attention has been directed towards capturing a representative sample of each feature.

@@

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets. One method to incorporate data on the ecological and evolutionary processes that sustain biological features is to conserve a representative sample of the physical attributes that underpin these processes. To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2]. However, using this method is challenging because biodiversity often cannot be divided into operational groups for conservation planning without loss of information [@r489; @r490; @r218]. This limitation has been known of quite some time, and dates back to some of the earliest reserve selection methods [@r489].

@@

The \texttt{raptr R} package utilizes two novel formulations of reserve selection problem to generate prioritizations. These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451]. Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the "unreliable" and "reliable" formulations. The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a given solution samples a feature's attribute space. On the other hand, the unreliable formulation assumes that features occupy all of the planning units within which they are found with 100 % certainty when performing these calculations. For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1. All mathematical terms defined hereafter are described in Table S1. For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.

@@

Targets are used to ensure that prioritizations sufficiently conserve each feature. Amount-based targets specify the minimum amount of habitat required for each feature to be adequately conserved (similar to those used in \texttt{Marxan}). Let $T_f$ denote the amount of area or habitat that needs to be preserved for each feature $f \in F$. Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature to be sufficiently conserved. Space-based targets are expressed as proportions--instead of a sum of weighted distances--by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid. This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means clustering analyses from the within sums of squares and total sums of squares [@r491; page 106]. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$. The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.

@@

## CASE STUDY 1
### Methods
We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the species' realized niches. By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience to environmental change [@r63]. We selected Queensland, Australia as the study area. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1). We used the scores from the first `r as.character(english::as.english(cs1_parameters$number_pc_layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1_pca_data[[4]][cs1_parameters$number_pc_layers], 1)` % of the total variation; Fig. 4).

We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale headed rosella (_Platycercus adscitus_). These species span a range of different habitat requirements. We mapped the extent of occurrence for each species (Fig. 5). To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455]. We used this method because it is entirely reproducible using freely available data.

@@

### Results
Generally, the solution generated using just amount targets preserved a representative sample of each the `r as.character(english::as.english(length(cs1_parameters$species_names)))` bird species' niches. (Figure S1). This solution secured over `r cs1_parameters$space_target*100` % of the realized niche of `r parsed_representative_space_held_names`. Yet it failed to achieve this for the `r parsed_not_representative_space_held_names`. This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.

@@

We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented using the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458]. These continuous variables described the main axes of genetic variation within the species. We averaged together values corresponding to samples collected in the same grid cell, and used them to create a genetic attribute space (Figs. 7c and 7d). To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using scaled  great circle distances between the grid cells' centroids [using the \texttt{ape R} package; @r464].

@@

The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2_spp_results$genetic[1], 2)` % sampled; Figs. 8a and 8c). This solution only conserved individuals in one of the two main genetic groups. In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target. Whereas, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs. 8b and 8d). Although the solutions only differ by a single planning unit, by securing individuals in both main genetic groups the second solution was able to secure a much more representative sample of the species' genetic variation for only a minor increase in cost (`r round(cs2_spp_results$Score[2],2)` compared to `r round(cs2_spp_results$Score[1],2)` total cost).

@@

The degree to which a prioritization truly secures a representative sample of a feature depends on (1) the attribute space(s), (2) the distribution of demand points chosen by the conservation planner, and (3) the space target used. The optimal solution will not effectively represent the decision maker's objectives if an inappropriate set of spatial variables or demand points are used to represent the attribute space. Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space (see the `make.DemandPoints` function for statistical routines to achieve this). As with conventional amount-based targets, planners will have to ensure that space targets are high enough to fulfill conservation objectives [eg. when setting targets to conserve intra-specific genetic variation; @r487]. We encourage conservation planners to consult with experts to identify suitable targets.

>>>>

Abstract word count: 344 / 350 \newline
Total word count: XXXX / 7000 \newline
Number of references: 42 \newline
Number of figures: 8 \newline
Number of tables: 0 \newline
Number of textboxes: 0 \newline
Number of equations: 4 \newline
Short running title: \texttt{raptr} R package \newline
Short abstract: Prioritize biodiversity processes and patterns using raptr https://github.com/jeffreyhanson/raptr  \newline
Keywords: conservation, biodiversity, mixed integer linear programming, optimization, protected areas  \newline

@@

1. An underlying aim in conservation planning is to maximize the long-term persistence of biodiversity. To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved. One way to conserve such processes at the feature level (eg. species, ecosystem) is to preserve a sample of the feature (eg. individuals, areas) that is representative in terms of the intrinsic or extrinsic physical attributes that underpin the process of interest. For example, by conserving a sample of populations with local adaptations--physical attributes associated with adaptation--that is representative of the range of adaptations found in the species, protected areas can maintain adaptive processes by ensuring these adaptations are not lost. Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of area or habitat for each feature. Little attention has been directed towards capturing a representative sample of the variation within each feature.

@@

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets. One method to incorporate data on the ecological and evolutionary processes that sustain biological features is to conserve a representative sample of the physical internal or external attributes that underpin these processes. To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2]. However, using this method is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490]. This limitation has been known of quite some time, and dates back to some of the earliest reserve selection methods [eg. @r295].

@@

The \texttt{raptr R} package utilizes two novel formulations of reserve selection problem to generate prioritizations. These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451]. Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the "unreliable" and "reliable" formulations. The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a given solution samples a feature's attribute space. On the other hand, the unreliable formulation assumes that features occupy all of the planning units within which they are found with absolute certainty when performing these calculations. For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1. All mathematical terms defined hereafter are described in Table S1. For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.

@@

Targets are used to ensure that prioritizations sufficiently conserve each feature. Amount-based targets specify the minimum amount of habitat required for each feature to be adequately conserved (similar to those used in \texttt{Marxan}). Let $T_f$ denote the amount of area or habitat that needs to be preserved for each feature $f \in F$. Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature to be sufficiently conserved. Space-based targets are expressed as proportions--instead of a sum of weighted distances--by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid. This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means clustering analyses from the within sums of squares and total sums of squares [page 106; @r491]. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$. The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.

@@

## CASE STUDY 1
### Methods
We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the species' realized niches. By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience to environmental change [@r63]. We selected Queensland, Australia as the study area, and used an equal area coordinate system for geosptial analyses (Australian Albers GDA94; \texttt{EPSG:3577}). We generated planning units by clipping a `r cs1_parameters$pu_size/1000` km$^2$ grid to the study area. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1). We used the scores from the first `r as.character(english::as.english(cs1_parameters$number_pc_layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1_pca_data[[4]][cs1_parameters$number_pc_layers], 1)` % of the total variation; Fig. 4).

We used four bird species in this case study: blue-winged kookaburra (_Dacelo leachii_), brown-backed honeyeater (_Ramsayornis modestus_), brown falcon (_Falco berigora_), and pale-headed rosella (_Platycercus adscitus_). These species span a range of different habitat requirements. We mapped the extent of occurrence for each species (Fig. 5). To do this, we obtained occurrence records from the Atlas of Living Australia across the whole of Australia [using the \texttt{ALA4R R} package; @r453], spatially thinned the data to omit points within `r cs1_parameters$thin_distance/1000` km$^2$ of each other to ameliorate the effects of sampling bias [using the developmental version of the \texttt{spThin R} package; \url{https://github.com/mlammens/spThin}; @r454], and fit `r cs1_parameters$mcp_percent` % minimum convex polygons [using the \texttt{adehabitatHR R} package; @r455]. We used this method because it is entirely reproducible using freely available data.

@@

### Results
Generally, the solution generated using just amount targets preserved a representative sample of each the `r as.character(english::as.english(length(cs1_parameters$species_names)))` bird species' niches (Figure S1). This solution secured over `r cs1_parameters$space_target*100` % of the realized niche of `r parsed_representative_space_held_names`. Yet it failed to achieve this for the `r parsed_not_representative_space_held_names`. This result demonstrates that although conventional methods may yield solutions that conserve a representative sample of some features, only through the use of explicit targets can planners obtain cost-effective solutions that secure a representative sample of all features.

@@

We used non-metric multi-dimensional scaling [NMDS; using Gower distances to accommodate sparsity; @r459; implemented using the \texttt{cluster R} package; @r457] to ordinate the presence (or absence) of locus-specific alleles within individuals into two continuous variables [implemented in the \texttt{vegan R} package; @r458]. These continuous variables described the main axes of genetic variation within the species. We averaged together values corresponding to samples collected in the same grid cell, and used them to create a genetic attribute space (Figs. 7c and 7d). To assess spatial auto-correlation, we calculated Moran's I auto-correlation index for each NMDS axis using inverse great circle distances between the grid cells' centroids as weights [using the \texttt{ape R} package; @r464].

@@

The solution generated using just the amount target failed to preserve a representative portion of the species' genetic variation (`r round(cs2_spp_results$genetic[1], 2)` % sampled; Figs. 8a and 8c). This solution only conserved individuals in one of the two main genetic groups. In fact, this solution is just a collection of the cheapest planning units occupied by the species needed to fullfil the target. Alternatively, the solution generated using both amount and genetic targets conserved individuals in each of the two main genetic groups (`r round(cs2_spp_results$genetic[2],2)` % sampled; Figs. 8b and 8d). Although the solutions only differ by a single planning unit, by securing individuals in both main genetic groups the second solution was able to secure a much more representative sample of the species' genetic variation for only a minor increase in cost (`r round(cs2_spp_results$Score[2],2)` compared to `r round(cs2_spp_results$Score[1],2)` total cost).

@@

The degree to which a prioritization truly secures a representative sample of a feature depends on (1) the attribute space(s), (2) the distribution of demand points chosen by the conservation planner, and (3) the space target used. The optimal solution will not effectively represent the decision maker's objectives if an inappropriate set of spatial variables or demand points are used to represent the attribute space. Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space (see the `make.DemandPoints` function for statistical routines to achieve this). As with conventional amount-based targets, planners will need ensure that space targets are high enough to fulfill conservation objectives [eg. when setting targets to conserve intra-specific genetic variation; @r487]. The most appropriate space target will depend on the species of conservation interest [eg. threatened species may more of their genetic variation to be conserved; @r495], and how conserving more variation relates to the species' persistence [eg. the relationship between inbreeding and extinction has a threshold; @r496]. We encourage conservation planners to consult with experts to identify suitable targets.

### 2017-04-24 code/rmarkdown/figures.Rmd
assorted fixes

---
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: preamble.tex
---

@@

```{r, echo = FALSE, fig.height = 1.9, fig.width = column_width, fig.cap = "Two main gradients of climatic variation across Queensland, Australia. Polygons denote planning units."}
# prepare data
cs1_pca_fpoints <- plyr::ldply(seq_len(cs1_parameters$number_pc_layers),
  function(i) {
    curr_data <- cs1_bioclim_raster[[i]] %>%
                     raster::aggregate(10) %>%
                     raster::rasterToPoints() %>%
                     as.data.frame()
    names(curr_data)[3] <- "Score"
    curr_data$pc <- paste0("PC ",i)
    return(curr_data)
})
# make plots
p1 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = aus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = "grey85", color = "grey70") +
  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
                                            pc == unique(pc)[1]),
                       ggplot2::aes(x = x, y = y, fill = Score)) +
  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = NA, color = "black", size = 0.05) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color = "transparent",
                                              fill = "transparent"),
    legend.position = c(1.1, 1.07),
    legend.justification = c(1,1),
    legend.key.height = ggplot2::unit(1.5, "mm"),
    legend.key.width = ggplot2::unit(2, "mm")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::scale_fill_continuous(breaks =
    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[1])$Score, n = 3,
          min.n = 2)) +
  ggplot2::facet_wrap(~ pc)
p2 <- ggplot2::ggplot() +
  ggplot2::geom_polygon(data = aus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill="grey85", color="grey70") +
  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
                                            pc == unique(pc)[2]),
                       ggplot2::aes(x = x, y = y, fill = Score)) +
  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
                        ggplot2::aes(x = long, y = lat, group = group),
                        fill = NA, color = "black", size = 0.05) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.line = ggplot2::element_blank(),
    plot.margin = ggplot2::unit(c(0, 0, 0, 0), "lines"),
    strip.background = ggplot2::element_rect(fill = "grey20"),
    strip.text = ggplot2::element_text(color = "white"),
    axis.title = ggplot2::element_blank(),
    legend.background = ggplot2::element_rect(color="transparent",
                                              fill="transparent"),
    legend.position = c(1.1, 1.07),
    legend.justification = c(1,1),
    legend.key.height = ggplot2::unit(1.5, "mm"),
    legend.key.width = ggplot2::unit(2, "mm")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::scale_fill_continuous(breaks =
    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[2])$Score, n = 2,
           min.n = 2)) +
  ggplot2::facet_wrap(~ pc)
# render plots
gridExtra::grid.arrange(p1, p2, ncol = 2)
```

```{r, echo = FALSE, fig.height = 3.8, fig.width = column_width, fig.cap = "Distribution of the species used in the first case study. Polygons denote planning units. Planning units occupied by a given species are shown in light blue."}
# prepare data
cs1_pus_occ_fpolygons <- plyr::ldply(
  seq_along(cs1_parameters$common_names),
  function(i) {
    curr_data <- cs1_pus_fpolygons
    curr_data$species <- cs1_parameters$common_names[i]
    curr_data$Occupancy <- "absent"
    curr_pus <- cs1_ru@data@pu.species.probabilities %>%
                dplyr::filter(species == i) %>%
                `[[`("pu")
    curr_data$Occupancy[which(curr_data$id %in% curr_pus)] <- "present"
    return(curr_data)
})

>>>>

---
title: "Figures"
output:
  rmarkdown::pdf_document:
    toc: false
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: preamble.tex
fontsize: 11pt
documentclass: article
subparagraph: true
bibliography: references.bib
csl: reference-style.csl
---

@@

```{r, echo = FALSE, fig.height = 1.9, fig.width = column_width, fig.cap = "Two main gradients of climatic variation across Queensland, Australia. Polygons denote planning units. The map is rendered in an equal-area coordinate system (\texttt{EPSG:3577})"}
# prepare data
cs1_pca_fpoints <- plyr::ldply(seq_len(cs1_parameters$number_pc_layers),
  function(i) {
    curr_data <- cs1_bioclim_raster[[i]] %>%
                     raster::aggregate(10) %>%
                     raster::rasterToPoints() %>%
                     as.data.frame()
    names(curr_data)[3] <- "Score"
    curr_data$pc <- paste0("PC ",i)
    return(curr_data)
})
  ggplot2::geom_polygon(data = aus_fpolygons,
  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
                                            pc == unique(pc)[1]),
                       ggplot2::aes(x = x, y = y, fill = Score)) +
  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
                        fill = NA, color = "black", size = 0.05) +
    legend.background = ggplot2::element_rect(color = "transparent",
                                              fill = "transparent"),
    legend.position = c(1.1, 1.07),
    legend.justification = c(1,1),
    legend.key.height = ggplot2::unit(1.5, "mm"),
    legend.key.width = ggplot2::unit(2, "mm")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::scale_fill_continuous(breaks =
    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[1])$Score, n = 3,
  ggplot2::facet_wrap(~ pc)
  ggplot2::geom_polygon(data = aus_fpolygons,
  ggplot2::geom_raster(data = dplyr::filter(cs1_pca_fpoints,
                                            pc == unique(pc)[2]),
                       ggplot2::aes(x = x, y = y, fill = Score)) +
  ggplot2::geom_polygon(data = cs1_pus_fpolygons,
                        fill = NA, color = "black", size = 0.05) +
    legend.background = ggplot2::element_rect(color="transparent",
                                              fill="transparent"),
    legend.position = c(1.1, 1.07),
    legend.justification = c(1,1),
    legend.key.height = ggplot2::unit(1.5, "mm"),
    legend.key.width = ggplot2::unit(2, "mm")) +
  ggplot2::coord_cartesian(xlim = buffered_range(cs1_pus@bbox[1, ], 0.025),
                           ylim = buffered_range(cs1_pus@bbox[2, ], 0.025),
                           expand = FALSE) +
  ggplot2::scale_fill_continuous(breaks =
    pretty(dplyr::filter(cs1_pca_fpoints, pc == unique(pc)[2])$Score, n = 2,
  ggplot2::facet_wrap(~ pc)
gridExtra::grid.arrange(p1, p2, ncol = 2)

```{r, echo = FALSE, fig.height = 3.8, fig.width = column_width, fig.cap = "Distribution of the species used in the first case study. See Fig. 4 caption for conventions. Planning units occupied by a given species are shown in light blue."}
# prepare data
cs1_pus_occ_fpolygons <- plyr::ldply(
  seq_along(cs1_parameters$common_names),
  function(i) {
    curr_data <- cs1_pus_fpolygons
    curr_data$species <- cs1_parameters$common_names[i]
    curr_data$Occupancy <- "absent"
    curr_pus <- cs1_ru@data@pu.species.probabilities %>%
                dplyr::filter(species == i) %>%
                `[[`("pu")
    curr_data$Occupancy[which(curr_data$id %in% curr_pus)] <- "present"
    return(curr_data)
})

### 2017-04-24 code/rmarkdown/supporting-information.Rmd
assorted fixes

We generated solutions for the conservation planning data sets using different formulations and parameters. Specifically, the solutions were generated using the unreliable and reliable formulation, and several boundary length multipliers (`r paste(unique(benchmark_parameters$blm), sep = "and")`). The solutions were generated using amount-based targets of `r benchmark_parameters$amount_target*100` % and space-based targets of `r benchmark_parameters$space_target*100`. To assess variation in solve times, we generated and solved `r benchmark_parameters$replicates` random data sets per combination of number of features and planning units. All problems were solved to within `r general_parameters$MIPGap * 100` % of optimality and were solved using a single thread on a system with 64 GB RAM and an Intel(R) Xeon (R) CPU (E5-2630 v3; 2.40 GHz).

@@

The benchmark analysis revealed that moderately-sized conservation problems can be solved in a relatively short period of time (Figure S2). Problems were solved much quicker when using the unreliable formulation than when using the reliable formulation. The solve times could be reduced by using more threads.  We were not able to determine how long problems containing over 100 planning units would take to solve using the reliable formulation because the system did not contain enough would take to solve because our computer system did not have enough RAM. Future research could involve implementing more advanced methods to solve the optimization problems [eg. as described in @r16; @r15].

>>>>

We generated solutions for the conservation planning data sets using different formulations and parameters. Specifically, the solutions were generated using the unreliable and reliable formulation, and several boundary length multipliers (`r paste(unique(benchmark_parameters$blm), sep = "and")`). The solutions were generated using amount-based targets of `r benchmark_parameters$amount_target*100` % and space-based targets of `r benchmark_parameters$space_target*100` %. To assess variation in solve times, we generated and solved `r benchmark_parameters$replicates` random data sets per combination of number of features and planning units. All problems were solved to within `r general_parameters$MIPGap * 100` % of optimality and were solved using a single thread on a system with 64 GB RAM and an Intel(R) Xeon (R) CPU (E5-2630 v3; 2.40 GHz).

@@

The benchmark analysis revealed that moderately-sized conservation problems can be solved in a relatively short period of time (Figure S2). Problems were solved much quicker when using the unreliable formulation than when using the reliable formulation. The solve times could be further reduced by using more threads. We were not able to determine how long problems bigger problems would take to solve using the reliable formulation because the system did not have enough RAM. Future research could involve implementing more advanced methods to solve the optimization problems [eg. as described in @r16; @r15].
